Solving Collective Spin Hamiltonians with Decoherence

### 2018-10-02 soc_squeezing/cumulant_expansions.tex
greatly simplified a derivation

  = m_\mu + n_\mu - \sum_\kappa\p{r_{\mu\kappa}+r_{\kappa\mu}},

@@

    { m_\mu - \sum_{\kappa<\nu} r_{\mu\kappa} \choose r_{\mu\nu} }
    { n_\mu - \sum_{\kappa<\nu} r_{\kappa\mu} \choose r_{\nu\mu} }}

@@

  &= \sp{\prod_\mu \p{m_\mu}_{\sum_\kappa r_{\mu\kappa}}
    \p{n_\mu}_{\sum_\kappa r_{\kappa\mu}}}

@@

Expanding the product
$\sigma_\mu\sigma_\nu=\sum_\kappa\eta_{\mu\nu\rho}\sigma_\rho$, where
unlike the sums over $\mu,\nu$ in most of the work above, the sum over
$\rho$ here includes an index for the identity operator
$\sigma_0\equiv\1$, we can expand and collect terms to get

@@

  = \prod_{\mu,\nu,a} \sum_\rho
  \eta_{\mu\nu\rho} \sigma_\rho^{(K^{\mu\nu}_a)}
  = \sum_{\v\rho} \prod_{\mu,\nu,a} \eta_{\mu\nu\rho^{\mu\nu}_a}
  \sigma_{\rho^{\mu\nu}_a}^{(K^{\mu\nu}_a)},

@@

where we have distributed a product of sums into a sum of products.
Collecting together operators of type $\sigma_\kappa$, we then have

@@

  = \sum_{\v\rho} \sp{\prod_{\mu,\nu,\kappa}
    \p{\eta_{\mu\nu\kappa}}^{\tilde\rho^{\mu\nu}_\kappa}}
  \prod_{\kappa,a} \sigma_{\kappa}^{(K^{\v\rho\kappa}_a)},

@@

where $\tilde\rho_\kappa^{\mu\nu}$ is the number of elements in

@@

equal to $\kappa$, and $K^{\v\rho\kappa}_a$ indexes the $a$-th
operator of type $\sigma_\kappa$ for each choice of $\v\rho$, such
that for any fixed $\v\rho$ we have
$\set{K^{\v\rho\kappa}_a}=\set{K^{\mu\nu}_b}$.  Exploiting the
exchange symmetry of $Q_{\v K}$ under the sum in \eqref{eq:SS_PQ}, we
can group together all terms which are equal up to a permutation of
indices in $\v\rho^{\mu\nu}$, i.e.~terms with equal values of
$\tilde\rho_\kappa^{\mu\nu}$, to write

@@

  \simeq \sum_{\tilde{\v\rho}}
  g_{\tilde{\v\rho}} \sp{\prod_{\mu,\nu,\kappa}
    \p{\eta_{\mu\nu\kappa}}^{\tilde\rho^{\mu\nu}_\kappa}}
  \prod_{\kappa,a} \sigma_{\kappa}^{(K^{\tilde{\v\rho}\kappa}_a)}
  = \sum_{\tilde{\v\rho}} g_{\tilde{\v\rho}}
  \sp{\prod_{\mu,\nu,\kappa}
    \p{\eta_{\mu\nu\kappa}}^{\tilde\rho^{\mu\nu}_\kappa}}
  P_{\v K_{\tilde{\v\rho}}},

@@

where $\simeq$ denotes equality up to a permutation of spin indices;
the degeneracy factor for each term is

@@

  \f{r_{\mu\nu}!}{\prod_\kappa\tilde\rho^{\mu\nu}_\kappa!};

@@

as before $K^{\tilde{\v\rho}\kappa}_a$ indexes the $a$-th operator of
type $\sigma_\kappa$ for each partitioning $\tilde{\v\rho}$, such that
for any fixed $\tilde{\v\rho}$ we have
$\set{K^{\tilde{\v\rho}\kappa}_a}=\set{K^{\mu\nu}_b}$; and
$\v K_{\tilde{\v\rho}}$ is an ordered list of the indices in
$\set{K^{\tilde{\v\rho}\kappa}_a}$.

% \begin{align}
%   \S_{\v m}\S_{\v n}
%   = \sum_{s\ge0} \sum_{\p{\v J,\v K,\v r;\v m,\v n,s}}
%   \sum_{\tilde{\v\rho}}
%   f_{\v m\v n\v r} g_{\tilde{\v\rho}}
%   \sp{\prod_{\mu,\nu,\kappa}
%     \p{\eta_{\mu\nu\kappa}}^{\tilde\rho^{\mu\nu}_\kappa}}
%   P_{\v J} P_{\v K_{\tilde{\v\rho}}}
% \end{align}

>>>>

  = m_\mu + n_\mu - \sum_\rho\p{r_{\mu\rho}+r_{\rho\mu}},

@@

    { m_\mu - \sum_{\rho<\nu} r_{\mu\rho} \choose r_{\mu\nu} }
    { n_\mu - \sum_{\rho<\nu} r_{\rho\mu} \choose r_{\nu\mu} }}

@@

  &= \sp{\prod_\mu \p{m_\mu}_{\sum_\rho r_{\mu\rho}}
    \p{n_\mu}_{\sum_\rho r_{\rho\mu}}}

@@

As the sum in \eqref{eq:SS_PQ} is invariant under permutation of the
indices in $\v K$, we can safely neglect keeping track of these
indices and simply write

@@

  = \prod_{\mu,\nu} \prod_{a=1}^{r_{\mu\nu}}
  \sigma_\mu^{(K^{\mu\nu}_a)} \sigma_\nu^{(K^{\mu\nu}_a)}
  \simeq \bigotimes_{\mu,\nu} \bigotimes_{a=1}^{r_{\mu\nu}}
  \sigma_\mu \sigma_\nu
  = \bigotimes_{\mu,\nu} \bigotimes_{a=1}^{r_{\mu\nu}}
  \sum_\rho \eta_{\mu\nu\rho} \sigma_\rho,

@@

where $\simeq$ denotes equality up to a re-indexing of spins; we have
introduced explicit dependence on
$r_{\mu\nu}=\abs{\set{K^{\mu\nu}_a}}$ for brevity; and
$\eta_{\mu\nu\rho}$ is defined by
$\sigma_\mu\sigma_\nu=\sum_\rho\eta_{\mu\nu\rho}\sigma_\rho$.  Unlike
the sums over $\mu,\nu$ in most of the work above, the sum over $\rho$
here includes an index for the identity operator $\sigma_0\equiv\1$.
Distributing the product of sums into a sum of products gives us

@@

  \simeq \sum_{\v\rho} \bigotimes_{\mu,\nu,a}
  \eta_{\mu\nu\rho^{\mu\nu}_a} \sigma_{\rho^{\mu\nu}_a}
  = \sum_{\v\rho} \sp{\prod_{\mu,\nu,a}
    \p{\eta_{\mu\nu\rho^{\mu\nu}_a}}^{\rho^{\mu\nu}_a}}
  \bigotimes_{\mu,\nu,a} \sigma_{\rho^{\mu\nu}_a}.
  \label{eq:Q_K_rho}

@@

Letting $\tilde\rho^{\mu\nu}_\kappa$ denote the number of elements in

@@

that are equal to $\kappa$, we observe that two terms in
\eqref{eq:Q_K_rho} with, say, $\v\rho=\v\rho_1$ and $\v\rho=\v\rho_2$
are equal up to a permutation of indices if
$\tilde{\v\rho_1}=\tilde{\v\rho_2}$.  We therefore collect such terms
together to write

@@

  \simeq \sum_{\tilde{\v\rho}} g_{\tilde{\v\rho}} \v\eta_{\tilde{\v\rho}}
  \bigotimes_{\kappa}
  \sigma_\kappa^{\otimes\sum_{\mu,\nu}\tilde\rho^{\mu\nu}_\kappa},
  \v\eta_{\tilde{\v\rho}}
  \p{\eta_{\mu\nu\kappa}}^{\tilde\rho^{\mu\nu}_\kappa},

@@

@@

  \f{r_{\mu\nu}!}{\prod_\kappa\tilde\rho^{\mu\nu}_\kappa!}

@@

counts the degeneracy of terms in \eqref{eq:Q_K_rho}, i.e.~the number
of $\v\rho$ which are consistent with $\tilde{\v\rho}$.
